Approximation as Differentiation

### 2025-07-10 main/language.tex
Some progress..

Models of $\Sigma$ can be given in any category $\cat{C}$ with finite products and a terminal object; such a
model $M$ assigns to each base type $\rho \in \PrimTy$ an object $\sem{\rho}_{M}$ in $\cat{C}$, and to each
primitive operation $\phi \in \PrimOp^\rho_{\rho_1,\ldots,\rho_n}$, a morphism $\sem{\phi}_{M}:
\sem{\rho_1}_{M} \times \ldots \times \sem{\rho_n}_{M} \to \sem{\rho}_{M}$.

To give the semantics for the full language, we fix a bicartesian closed category $\Sem$ plus a model $M$ of
$\Sigma$ in $\Sem$, with the following additional structure:
\begin{enumerate}
\item a strong monad $(\Lift, \eta, \mu)$ with strength $\mathsf{st}_{X,Y}: X \times \Lift(Y) \to \Lift(X
\times Y)$
\item an endofunctor $\List: \Sem \to \Sem$, plus for any object $X$, morphisms $\nil: 1 \to \List(X)$ and
$\cons: X \times \List(X) \to \List(X)$, and for any objects $Y, Z$ and morphisms $f_\nil: Z \to Y$ and
$f_\cons: Z \times X \times Y \to Y$, a morphism $\fold(f_\nil,f_\cons): Z \times \List(X) \to Y$.
\end{enumerate}

>>>>

A model $M$ of $\Sigma$ can be given in any category $\cat{C}$ with finite products and a terminal object, and
assigns to each base type $\rho \in \PrimTy$ an object $\sem{\rho}_{M}$ in $\cat{C}$, and to each primitive
operation $\phi \in \PrimOp^\rho_{\rho_1,\ldots,\rho_n}$, a morphism $\sem{\phi}_{M}: \sem{\rho_1}_{M} \times
\ldots \times \sem{\rho_n}_{M} \to \sem{\rho}_{M}$. A model of $\tyList$ can also be given in $\cat{C}$ and
consists of an endofunctor $\List: \cat{C} \to \cat{C}$, plus for any object $X$, morphisms $\nil: 1 \to
\List(X)$ and $\cons: X \times \List(X) \to \List(X)$, and for any objects $Y, Z$ and morphisms $f_\nil: Z \to
Y$ and $f_\cons: Z \times X \times Y \to Y$, a morphism $\fold(f_\nil,f_\cons): Z \times \List(X) \to Y$.

To give the semantics for the full language, we fix a bicartesian closed category $\Sem$ with a strong monad
$(\Lift, \eta, \mu)$ with strength $\mathsf{st}$, a model $M$ of $\Sigma$ in $\Sem$, and a model of $\tyList$
in $\Sem$.
